Using Machine Learning and Constraint Programming to Solve the Relational-To-Ontology Mapping Problem

### 2017-08-11 stp/paper_aaai/main.tex
moved some paragraphs around, writtten pattern mining part

\usepackage{tikz}
\usetikzlibrary{shadows.blur}
\usetikzlibrary{shapes.symbols}
\usetikzlibrary{arrows,shapes,decorations.pathmorphing}
\usetikzlibrary{positioning}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzstyle{class}=[draw,ellipse,align=center,fill=green!25]
\tikzstyle{data}=[draw,rectangle,align=center,black!70]
\tikzstyle{hasaA} = 
[>=o,draw,densely dashed,-,black!70,decoration={snake,segment 
length=10,amplitude=1.5,post length=4}]
\tikzstyle{hasaC} = [>=stealth',draw,thick,-, black]
\tikzstyle{isa} = [>=latex,draw,double,black!90,double distance=1mm]

@@

%  \pdfinfo{
%/Title (2018 Formatting Instructions for Authors Using LaTeX)
%/Author (AAAI Press Staff)}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Using Machine Learning and Constraint Programming to Solve the 
Relational-To-Ontology Mapping Problem}
\author{}
\maketitle
\begin{abstract}
The problem of integrating heterogeneous data sources in an ontology is of high 
relevance in the database field. Several techniques exist to approach this 
problem. Nonetheless, side constraints on the data integration cannot be easily 
integrated with these tools and thus the results may be inconsistent. We 
present a new approach that combines Machine Learning and 
Constraint Programming techniques, by modelling the data integration problem as 
a Steiner Tree Problem in a weighted graph. We show that through this approach 
we can achieve better precision, recall and speed compared to state-of-the-art 
approaches. We provide a comprehensive set of experiments supporting our 
findings.
\end{abstract}

@@

In our case, we would like this global model to account
not only for structural properties of the original data sources, but also to 
include the semantics which is usually implicitly present in the sources. In 
other words, we want to build a semantic model which describes the data sources 
in terms of concepts and relationships defined by an ontology. We can view this
semantic model as a graph with ontology classes as the nodes and ontology
properties as the links between the nodes \cite{Taheriyan2013}. Henceforth, we 
focus on a specific data integration problem: automatically mapping relational 
data sources onto a user provided ontology.

In this paper we use Machine Learning techniques (ML) to learn mapping rules 
from already mapped instances. To this end we formulate the 
Relational-To-Ontology Mapping Problem (\relonto{})
as a Minimum Cost Steiner Tree Problem (STP) with side constraints.
This problem is formulated for an \emph{integration graph} which includes the 
attributes from the new source as well as ontology classes and properties. We
apply machine learning techniques to assign costs to the links in the 
constructed integration graph.

@@

\end{tikzpicture}
\caption{Example of ontology
(`\tikz{ \node[class] () at (0,0)  {};}' are Classes, `\tikz{ \path[hasaC,->] 
(0,1) -- (3ex,1);}' means ``has a Class property'', 
`\tikz{ \path[hasaA] (0,1) edge[decorate] (3ex,1);}' means ``has a data 
property'' and 
`\tikz{ \path[isa,->] (0,1) -- (3ex,1);}' means ``is a'' ).
}
\label{FIG:onto}
\end{figure}

@@

\section{ML for Training on Source Descriptions \ddg{Better title than this 
please...}}

@@

\ddg{TODO}

@@

The Steiner Tree Problem (STP) is defined the following way. Given a graph $G =
(V, E)$ and a subset of its nodes $T \subseteq V$ , called \emph{terminals}, a 
Steiner Tree $G_s = (V_s, E_s )$ is a tree such that $T \subseteq V_s \subseteq 
V$ and $E_s \subseteq E$. In other words, $G_s$ spans all the nodes in $T$ and 
may include additional nodes from $V$, in particular to ensure the 
connectedness of the constructed tree.

@@

Note that, although the alignment graph is directed, the semantic models are 
not necessarily rooted directed trees, therefore the directions of the edges is 
ignored for the STP model (and are all treated as undirected edges).

@@

We need to build a subgraph $T^\star= (V^\star, E^\star)$ 
that is a subset of the integration graph 
$\mathcal{I}_\mathcal{O}^{s^\star}$ for the new source $s^\star$. The tree will 
be used to build the source description $\delta^\star = (s^\star, m^\star, 
\phi^\star)$. 
In particular, $(\mathcal{V_O} \cap V^\star,\mathcal{E_O} \cap E^\star)$ 
corresponds to the semantic model, and 
$(\mathcal{A}_{s^\star},\mathcal{M}_\mathcal{O}^{s^\star} \cap E^\star)$ 
corresponds to 
the attribute mapping function.

The solution $T^\star$ must satisfy the following constraints:
\begin{enumerate}
	\item $T^\star$ must be a tree (connected acyclic graph).
	\item $\forall a \in \mathcal{A}_{s^\star}, a\in V^\star$.
	\item $\forall a \in \mathcal{A}_{s^\star}, degree(a) = 1$.
	\item $\forall n \in \mathcal{D_{G_O}}, degree(a) \in 
	\{0,2\}$
	\item $\forall n \in \mathcal{D_{G_O}} \cap V^\star, 
	degree(a) = 2$
\end{enumerate}

>>>>

\usepackage{tikz}
\usetikzlibrary{shadows.blur}
\usetikzlibrary{shapes.symbols}
\usetikzlibrary{arrows,shapes,decorations.pathmorphing}
\usetikzlibrary{positioning}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzstyle{class}=[draw,ellipse,align=center,fill=green!25]
\tikzstyle{data}=[draw,rectangle,align=center,black!70]
\tikzstyle{hasaA} = 
[draw,densely dashed,-,black!70,decoration={snake,segment 
length=10,amplitude=1.5,post length=4}]
\tikzstyle{hasaC} = [>=stealth',draw,thick,-, black]
\tikzstyle{isa} = [>=latex,draw,double,black!90,double distance=1mm]

@@

%  \pdfinfo{
%/Title (2018 Formatting Instructions for Authors Using LaTeX)
%/Author (AAAI Press Staff)}
\setcounter{secnumdepth}{2}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Using Machine Learning and Constraint Programming to Solve the 
Relational-To-Ontology Mapping Problem}
\author{}
\maketitle
\begin{abstract}
The problem of integrating heterogeneous data sources in an ontology is of high 
relevance in the database field. Several techniques exist to approach this 
problem. Nonetheless, side constraints on the data integration cannot be easily 
integrated with these tools and thus the results may be inconsistent. We 
present a new approach that combines Machine Learning and 
Constraint Programming techniques, by modelling the data integration problem as 
a Steiner Tree Problem in a weighted graph. We show that through this approach 
we can achieve better precision, recall and speed compared to state-of-the-art 
approaches. We provide a comprehensive set of experiments supporting our 
findings.
\end{abstract}

@@

In our case, we would like this global model to account
not only for structural properties of the original data sources, but also to 
include the semantics which is usually implicitly present in the sources. In 
other words, we want to build a semantic model which describes the data sources 
in terms of concepts and relationships defined by an ontology. We can view this
semantic model as a graph with ontology classes as the nodes and ontology
properties as the links between the nodes \cite{Taheriyan2013}. Henceforth, we 
focus on a specific data integration problem: automatically mapping a new 
relational data source onto a user provided ontology.

In this paper we use Machine Learning techniques (ML) to learn mapping rules 
from already mapped instances. To this end we formulate the 
Relational-To-Ontology Mapping Problem (\relonto{})
as a Minimum Cost Steiner Tree Problem (STP) with side constraints 
\cite{deuna2016steiner}.

Firstly, we build an \emph{integration graph} which includes the 
attributes from the new source as well as ontology classes and properties. 
Secondly, we apply machine learning techniques to 
assign costs to its edges. 
Lastly, we use Constraint Programming to 
find a minimum cost Steiner Tree in the graph. The goal is to assign the costs 
on the edges in such a way that the resulting Steiner Tree is a valid semantic 
model for the new source.

@@

\end{tikzpicture}
\caption{Example of ontology
(`\tikz{ \node[class] () at (0,0)  {};}' are Classes, `\tikz{ \path[hasaC,->] 
(0,1) -- (3ex,1);}' means ``has a Class property'', 
`\tikz{ \path[hasaA,decoration={post length=0}] (0,0) edge[decorate] (3ex,0);}' 
means ``has 
a data 
property'' and 
`\tikz{ \path[isa,->] (0,1) -- (3ex,1);}' means ``is a'' ).
}
\label{FIG:onto}
\end{figure}

@@

\section{ML for Training on Source Descriptions \label{SEC:ML}\ddg{Better title 
than this 
please...}}

@@

As an extension to the problem, we consider using pattern mining techniques in 
order to identify patterns in the training set of semantic model that repeat 
themselves often. The hope is that with such information we can encourage the 
semantic model $m^\star$ for the new source to use groups of edges that 
appeared often together in training semantic models.

In our context, we define a \emph{pattern} is a graph $P$ that is undirected 
and simple (no multiple edges between two nodes, nor self-loops).

Let $G = (V_G,E_G,\lambda_G)$ and $P = (V_P,E_P, \lambda_P)$ be two graphs 
where $\lambda_P$ is a \emph{labeling function} that maps nodes/edges of $G$ 
into a set of labels (e.g. natural numbers) $\mathbb{N}$ (similarly for 
$\lambda_P$).
An \emph{embedding} of $P$ into $G$ is an injective function $f : V_G \mapsto 
V_P$ such that for all $x,y \in V_P$:
\begin{enumerate}
	\item $\{x,y\} \in E_P \implies \{f(x),f{y}\} \in E_G$.
	\item $\lambda_P(x) = \lambda_G(f(x))$
	\item $\lambda_P(\{x,y\}) = \lambda_G(\{f(x),f(y)\})$
\end{enumerate} 
That is, if an edge exists in $P$, it must also exist in $G$ (but not 
vice-versa); and the labelling functions of both graphs must match after 
applying the embedding function.

Our goal is to take a set of graphs (the semantic models from the training set 
$S_T$) and find patterns that appear often across that set of graphs. This is 
known as the Frequent Subgraph Mining (FSM), which contains the subgraph 
isomorphism, known to be NP-complete. 
There exists a range of tools that can achieve this task using Transactional 
Frequent Graph Pattern Mining
\cite{petermann2017dimspan,yan2002gspan}. We chose to use DIMSpan 
\cite{petermann2017dimspan}. \ddg{CHECK! How do they solve an NP-complete 
problem??}

This tool provides us with a list of patterns that are frequent, as well as 
their support. The \emph{support} of a pattern is a measure of frequency. Thus 
the higher the support, the more often the pattern appears. This typically 
means that big patterns have low support.

@@

Given a graph $G =
(V, E)$ and a subset of its nodes $T \subseteq V$ , called \emph{terminals}, a 
Steiner Tree $G_s = (V_s, E_s )$ is a tree such that $T \subseteq V_s \subseteq 
V$ and $E_s \subseteq E$. In other words, $G_s$ spans all the nodes in $T$ and 
may include additional nodes from $V$, in particular to ensure the 
connectedness of the constructed tree. The Steiner Tree Problem 
(STP) is stated as follows: given $G$ and a weight function $w_f : E 
\longmapsto \mathbb{R}$, find the Steiner Tree that minimizes the sum of the 
weights of the edges in $E_s$ given by $w_f$. This was proven to be NP-hard 
by Karp \cite{Karp1972}.

@@

Note that, although the alignment graph is directed, the semantic models are 
not necessarily rooted directed trees (e.g. \ref{FIG:sem}), therefore the 
directions of the edges is ignored for the STP model (and are all treated as 
undirected edges).

The goal is to build a subgraph $T^\star= (V^\star, E^\star)$ 
of the integration graph 
$\mathcal{I}_\mathcal{O}^{s^\star}$ for the new source $s^\star$. The solution 
$T^\star$ will 
be used to build the source description $\delta^\star = (s^\star, m^\star, 
\phi^\star)$. 
In particular, $(\mathcal{V_O} \cap V^\star,\mathcal{E_O} \cap E^\star)$ 
corresponds to the semantic model, and 
$(\mathcal{A}_{s^\star},\mathcal{M}_\mathcal{O}^{s^\star} \cap E^\star)$ 
corresponds to 
the attribute mapping function.

The solution $T^\star$ must satisfy the following constraints:
\begin{enumerate}
	\item $T^\star$ must be a subgraph of $\mathcal{I}_\mathcal{O}^{s^\star}$.
	\item $T^\star$ must be a tree (connected acyclic graph).
	\item $\forall a \in \mathcal{A}_{s^\star}, a\in V^\star$.
	\item $\forall a \in \mathcal{A}_{s^\star}, degree(a) = 1$.
	\item $\forall n \in \mathcal{D_{G_O}}, degree(a) \in 
	\{0,2\}$
	\item $\forall n \in \mathcal{D_{G_O}} \cap V^\star, 
	degree(a) = 2$
\end{enumerate}

It is therefore natural to model this a STP with side constraints, since the 2 
first requirements imply that $T^\star$ is a Steiner Tree of the integration 
graph. By designing the weighting function $w_\mathcal{I}$ through ML 
techniques, as shown in Section \ref{SEC:ML}, 
our hope is that the minimum weight Steiner Tree is a valid semantic model for 
the new source.

@@
